---
title: "Salicylate Statistics"
author: "Aaron Wong"
date: "15 September, 2022"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(lme4)
library(nlme)
library(ggplot2)
if(!require(here)){
  install.packages("here")
}
library(plyr)
if(!require(multcomp)){
  install.packages("multcomp")
}

a2dB <- function(a) {
  dB <- 20*log10(a)
  return(dB)
}

```

## Load Data

```{r getdata}
dat <- read.csv(here::here("gen/statistics/input/SalicylateThresholds.txt"),header=TRUE)
str(dat)
xtabs(~ batch + subject, dat, sparse = TRUE)

threshold <- dat$threshold
frequency <- factor(dat$frequency)
frequency <- relevel(frequency,ref="16") # set reference to 16 Hz
intensity <- factor(dat$intensity)
intensity <- relevel(intensity,ref="60") # set reference to 60 dB
condition <- factor(dat$condition)
batch <- factor(dat$batch)
subject <- factor(dat$subject)

```


##  Linear Mixed Effect Model

The following codes run a linear mixed effect model over the data. We will use the `lme` function from the `nlme` package which gives us p-values

```{r linear mixed model 2, echo=TRUE}
frequency <- relevel(frequency,ref="16") # set reference to 16 Hz
intensity <- relevel(intensity,ref="60") # set reference to 60 dB
SalLMM0 <- lme(threshold ~ intensity * frequency * condition, 
            random = ~ 1|batch/subject)
summary(SalLMM0)   # show results of linear model fit
```
```{r echo=FALSE}
#anova(SalLMM0)
#anova(SalLMM0, Terms = c("frequency", "frequency:condition"))
#anova(SalLMM0, Terms = c("condition", "frequency:condition"))

#plot(SalLMM0)


```
From the results above one can see that `conditionSalicylate` has a value of -5.01 +/- 1.28 (dB) and a p-value of 0.0003, while `conditionWashout` has a value of 1.00 +/- 1.28 (dB) and a p-value of 0.43.

The following re-run the Linear Mixed-Effect Model using 512 Hz as the reference modulation frequency. Thus the coefficient `conditionSalicylate` directly test whether 512 Hz salicylate at 60 dB is significantly different from baseline, and `conditionWashout` directly test that for washout vs baseline.

```{r releveling}
frequency <- relevel(frequency,ref="512") # set reference to 16 Hz
intensity <- relevel(intensity,ref="60") # set reference to 60 dB

SalLMM1 <- lme(threshold ~ intensity * frequency * condition, 
            random = ~ 1|batch/subject)
summary(SalLMM1)   # show results of linear model fit

```
From the results above one can see that `conditionSalicylate` has a value of -3.36 +/- 1.28 (dB) and a p-value of 0.011, while `conditionWashout` has a value of -0.28 +/- 1.28 (dB) and a p-value of 0.82.

### General Linear Hypothesis Testing

To create a more general environment for hypothesis testing, we will create an LMM `SalLMM2` using the package ``lme4`. This package does not give p-values when passed to `summary()`, but allows specific hypotheses to be tested using the package `glht` (General Linear Hypothesis Testing). We will relevel the frequency to use 16 Hz as the reference.

```{r linear mixed model, echo=TRUE}
frequency <- relevel(frequency,ref="16") # set reference to 16 Hz
intensity <- relevel(intensity,ref="60") # set reference to 60 dB
SalLMM2 <- lme4::lmer(threshold ~ frequency*intensity*condition + (1|batch/subject))
```

The following lines specify the hypotheses to be tested. Namely, we will test the following

1. Salicylate effect at 16 Hz, 60 dB
2. Salicylate effect at 512 Hz, 60 dB
3. Washout effect at 16 Hz, 60 dB
4. Washout effect at 512 Hz, 60 dB
5. Iceberg effect at 16 Hz
6. Iceberg effect at 512 Hz

```{r specific hypotheses}
tests <- c("conditionSalicylate = 0", # Salicylate effect for 16 Hz, 60 dB (ref-levels)
          "conditionSalicylate + frequency512:conditionSalicylate = 0", # Salicylate effect for 512 Hz, 60 dB
          "conditionWashout = 0", # Washout effect for 16 Hz, 60 dB (ref-levels)
          "conditionWashout + frequency512:conditionWashout = 0", # Washout effect for 512 Hz, 60 dB
          "conditionSalicylate - intensity45 = 0", #iceberg for 16Hz
          "conditionSalicylate + frequency512:conditionSalicylate - intensity45 - frequency512:intensity45 = 0") #iceberg for 512Hz
summary(glht(SalLMM2, linfct = tests))
```
#### Salicylate effects
Hypotheses 1,2 (salicylate effects) are significant, indicating that salicylate changes the AM threshold for both frequencies (at 60 dB).

#### Washout effects
Hypotheses 3,4 (Washout effects) are not significant, indicating that the AM threshold at washout period are not different from baseline (at 60 dB).

#### Iceberg effects
Hypothesis 5 compares the 16 Hz 60 dB salicylate to 16 Hz 45 dB baseline. On average this gives a -5.94 +/- 1.28 dB improvement in AM thresholdm, which is statistically significant, indicating that a 15 dB threshold shift cannot compensate for the change in AM threshold at 16 Hz upon salicylate. 

Hypothesis 6 compares the 512 Hz 60 dB salicylate to 512Hz 45 dB baseline. On average this gives a -2.55 +/- 1.28 dB improvement in AM threshold, which is slightly less than the -3.36 dB improvement when looking at 60 dB alone (Hypothesis 2). However, this was not statistically significant, suggesting that the AM threshold shift at 512 Hz could potentially be compensated by a 15 dB intensity threshold shift.
